---
title: "Running a resilient classroom brainwriting session"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running a resilient classroom brainwriting session}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(brainwritR)
```

## Prepare a useful activity

brainwritR supports a classroom adaptation of the 6-3-5 method attributed to
Bernd Rohrbach. The original name refers to six people, three ideas and five
minutes; this application instead provides **two questions per topic** and
rotates topics between groups. Its 15-2-5 preset is intended for a room of
approximately 15 people. It does not enforce a fixed number of ideas per answer.

Before the class, choose three topics that benefit from elaboration rather than
a single correct answer. For each, pair a generative question with a practical
follow-up. For example, a topic about teamwork could ask “Was erleichtert die
Zusammenarbeit?” and “Wie koennen wir das im naechsten Projekt umsetzen?”.
Explain that participants should read, extend, and combine earlier ideas rather
than judge their authors. Reserve time after the writing rounds to discuss the
results; the software organizes contributions but does not rank them.

Use pseudonyms. Explain who can see earlier drafts, who will receive exports,
and when records will be deleted. Participants need a modern browser, a reachable
server, and the same browser/origin throughout the session to preserve resume.
No participant account is created. One person should use one browser tab.

## Launch and configure

```{r launch, eval=FALSE}
run_app()
# Participant: http://localhost:3838
# Moderator:   http://localhost:3838/?mod=1
```

The default PIN `635` is convenient for a local demonstration. For a reachable
classroom instance, provide a private PIN and the real participant URL:

```{r classroom, eval=FALSE}
run_app(
  db_path = "/srv/brainwriting/data/classroom.sqlite",
  mod_pin = Sys.getenv("MOD_PIN"),
  base_url = "https://brainwriting.example.de",
  port = 3838,
  host = "0.0.0.0",
  poll_ms = 2500
)
```

`DB_PATH`, `MOD_PIN`, `BASE_URL`, and `PORT` supply defaults when their respective
arguments are omitted. Arguments win over environment variables. The QR code
uses `base_url` exactly: it must be reachable from phones and must not contain
`?mod=1`. The server does not infer its external hostname from proxy headers.
The default SQLite path is relative to the working directory; choose a stable
absolute path in an operational service.

The app creates the parent database directory at startup and initializes the
schema if necessary. Loading the R package alone has no filesystem effects.
The moderator PIN is held in process configuration, not in the session tables.

## Facilitate the four lifecycle stages

### Setup

The moderator opens `/?mod=1`, enters the PIN and supplies all topic titles and
both questions. There must be 2–6 topics and an equal number of groups. The setup
form defaults to three rounds of 300 seconds. Choose K rounds if every group
should encounter every topic exactly once. Fewer rounds produce a partial cycle;
additional rounds repeat topics.

### Lobby

Select **Speichern und Lobby oeffnen**. Project the QR code or share the URL.
Each participant enters a name or pseudonym and selects **Teilnehmen**. A blank
name receives an automatically generated label. Participants see a waiting view;
the moderator sees the live roster. Start requires at least K participants.

Groups are randomized and balanced at start, independent of joining order.
For 16 participants and K = 3 the group sizes are 6, 5, and 5. The corresponding
topics permanently receive 6, 5, and 5 sheets. Refreshing or reconnecting a browser
with a valid saved identity does not create a new participant.

### Running

Each participant sees a round/group/sheet label, topic title, shared clock, and
two answer fields. Previous rounds' nonempty contributions on the assigned sheet
appear above the corresponding question. Earlier **drafts as well as submitted
answers** are visible. Same-round contributions by another participant sharing
a sheet are kept independently, and are visible in later rounds and exports.

**Abgeben** submits both fields. Submission indicates progress, not a lock:
a participant may continue to edit, and an autosave preserves the submitted
flag. The moderator sees the number of submitted participants per group, can add
60 seconds, and can force the next round with **Runde beenden**. Early advancement
can cut off unsaved typing; announce it before using the control.

Late arrivals join the smallest group, using its next free index. They cannot
recover missed topics automatically, and do not create new sheets.

### Finished

After the last round the moderator sees grouped results and CSV/Markdown
buttons. Participants see a completion message. Export before selecting
**Neue Session …** and confirming **Ja, alles loeschen**. Cancel leaves the data
intact. Reset is accepted only for a finished session and removes all participants,
topics, contributions and assignments together. Connected clients discard their
old identity; a later returning browser also has its stale identity cleared.

## Understand the rotation invariant

```{r rotation-table}
k <- 3L
rotation <- outer(seq_len(k), seq_len(k), function(g, r) topic_for(g, r, k))
dimnames(rotation) <- list(paste("Group", 1:k), paste("Round", 1:k))
knitr::kable(rotation)
```

Every row and every column is a permutation of topics 1 through K. The formulas
are vectorized and exposed independently of the app:

```{r uneven}
sheet_for(idx = 1:6, n_sheets = 5)
sheet_for(idx = 1:5, n_sheets = 6)
```

In the first case two people share sheet 1 and each gets a distinct database row.
In the second case sheet 6 receives no contribution in that round. The initial
reference commentary claimed that every sheet receives a new group's entry
every round; that claim cannot hold for unequal groups under the specified
modulo mapping. The mapping is preserved and this limitation is made explicit.
The all-topics guarantee also requires attendance for the entire cycle.

## Persistence and timing contracts

SQLite contains four tables:

| Table | Purpose |
|:--|:--|
| `session` | Singleton state, configuration, current round, epoch deadline |
| `topics` | Titles, two questions, frozen sheet counts |
| `participants` | Opaque identity, pseudonym, group/index and joining time |
| `entries` | Topic/sheet/round/question/person text, submission flag, edit time |

The uniqueness key includes the participant ID. Multiple contributors sharing a
sheet never replace one another's entries. Each operation opens a connection,
sets a 5000 ms busy timeout, and closes it. WAL mode is enabled at initialization.
Setup, group assignment, joining and reset use transactions so a failure cannot
leave a half-applied lifecycle operation.

The round deadline is an epoch timestamp in SQLite. Both polling streams attempt
an advance, and an SQL guard on status and current round makes a racing update
idempotent. A new round's deadline is the time of advancement plus its duration.
There is no background scheduler without connected sessions: if everyone closes
the app, an expired round advances on the next connection, with a full duration
for the new round. The application does not replay multiple missed rounds.

Two polls separate metadata from contributions. Gated main views depend only
on real transitions; history, timer, roster and progress update independently.
Saved text is read without creating a reactive dependency during a main-view
render. Autosave captures the editing context with the input and writes after
1200 ms of quiet. This prevents a delayed previous-round edit from being written
into the next round. A late write may nevertheless appear only when a subsequent
view of that sheet is rendered.

### Mobile recovery and limits

On `shiny:disconnected`, the browser reloads after 1.5 seconds. On every
`shiny:connected`, it sends the locally stored participant ID, which the server
validates against SQLite. A valid identity resumes saved work. If local storage
is blocked or erased, automatic resume cannot work. Do not share a browser
profile between simultaneous participants.

The countdown redraws once per second; polling defaults to every 2.5 seconds.
It turns red below one minute. Briefly showing `0:00` before the next poll is
expected. Unsaved keystrokes at the boundary or while offline can be lost.
Submission before expiry is the clearest way to preserve a completed response.
There is no offline editing queue or merge mechanism for concurrent tabs.

## Deploy one durable instance

The repository includes `docker/Dockerfile` and `docker/docker-compose.yml`.
Build with the repository root as context. Compose expects an existing Traefik
installation, the `proxy` Docker network, a `websecure` entrypoint, a working
certificate resolver, and DNS for the public hostname. It does not provision
those shared infrastructure components.

```{bash deploy, eval=FALSE}
mkdir -p docker/data
sudo chown 10001:10001 docker/data
export BW_HOST=brainwriting.example.de
export TRAEFIK_CERTRESOLVER=letsencrypt
export MOD_PIN='replace-with-a-private-pin'
docker compose -f docker/docker-compose.yml up -d --build
```

The image runs as UID 10001. Bind-mounted `docker/data/` must be writable by that
UID. The SQLite file, WAL, and shared-memory files all belong on the persistent
volume. Do not mount only the `.sqlite` file. Use local disk and exactly one
application process. No replicas, ShinyProxy, or horizontal scaling are supported.
For another course, use another database/container and distinct Traefik labels
and hostname; do not share a database across instances.

For backups, stop the app before copying the entire database directory, or use
SQLite's online backup API. Copying a live `.sqlite` alone can omit committed
WAL data. Protect backups like the live database. Restarting a container with
the same volume resumes the same session; rebuilding the image is not a reset.
Keep the host clock synchronized since deadlines use wall-clock epoch seconds.

## Debrief and handle exports

CSV includes every saved entry, even an empty draft, and the exact German columns
`Thema`, `Bogen`, `Runde`, `Frage`, `Teilnehmer`, `Beitrag`, `abgegeben`, `Zeit`.
The final column uses the server's local timezone. Compose sets `Europe/Berlin`.
Markdown includes topic questions and nonempty entries grouped by topic and sheet.
Both formats include submitted and unsubmitted contributions.

Use the protocol to identify themes together. Participant names and contribution
text remain untrusted input in exports: import CSV columns as text to avoid
spreadsheet formula execution, and render Markdown using a safe renderer that
disables raw HTML. The interactive app renders user text as escaped content.

Self-hosting and pseudonyms reduce exposure but do not establish legal compliance
by themselves. Provide a suitable notice and retention policy for the course.
The reset action removes logical records, not old exports, backups, free SQLite
pages or filesystem snapshots. Separate retention and disposal procedures apply.

## Troubleshooting

| Symptom | Check |
|:--|:--|
| Phone cannot open QR link | Public `BASE_URL`, DNS, Wi-Fi reachability and TLS |
| App cannot create database | Directory ownership for UID 10001 and mounted volume |
| Grey screen/reload loop | Reverse proxy WebSocket support and service availability |
| Participant appears twice | Cleared storage, another browser, or an earlier separate join |
| Timer paused while room was absent | Expected: no clients means no polling scheduler |
| Some sheets lack a round | Unequal group sizes, late arrival or unanswered questions |
| Old work appears after restart | Expected persistence; finish/export/reset for a new course |

## Validation and reference changes

The archived single-file draft was the refactoring reference. The table schema,
rotation formulas, upsert semantics and guarded advancement SQL were retained.
Deliberate corrections include server-side checks on all moderator operations,
transactional lifecycle writes, stale-identity cleanup, avoiding premature
identity removal after joining, preserving the context of debounced edits, and
preventing a delayed post-reset write from recreating deleted entries. The
finished participant route refuses new joining instead of presenting a futile
join form. CSV column capitalization follows the requested export contract.
The viewport permits zoom and inputs have accessible labels.

Tests exercise these changes rather than asserting only successful page startup.
Run the package tests, browser suite, package check, lint and documentation build
before deploying a changed version. A passing automated browser suite does not
replace a classroom rehearsal on the institution's actual phones and network.
