---
title: "Running and reviewing a classroom brainwriting session"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running and reviewing a classroom brainwriting session}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(brainwritR)
```

## Plan the activity

brainwritR implements a classroom adaptation of the 6-3-5 method attributed to
Bernd Rohrbach. The original name refers to six people, three ideas and five
minutes. This application instead provides **two questions per topic** and
rotates topics between groups. The course variant 15-2-5 means about 15
participants, two questions and five minutes per round. In setup, enter the planned
number of participants: the form then names the resulting format, such as
"Format 15-2-5", with the group size and how many topics everyone will see.
The app does not enforce a fixed number of ideas in an answer. Names in setup are
optional; participants can enter a name or a pseudonym when they join.

Keep the core of 6-3-5 within the two fields: frame each topic as one open
problem, ask for new ideas in the first field and for developing an idea already
on the sheet in the second. For example: “How might we get everyone to take an
active part in group work? Note up to three new ideas.” and “Pick up an idea from
above (or one of your own) and develop it further.” Ask participants to read,
extend and combine earlier ideas. Reserve time after writing for discussion;
the application organizes contributions without ranking their quality.

Use pseudonyms. Explain who can read drafts, who will receive exports and when
records will be removed. A modern browser and reachable server are required.
There are no participant accounts. The interface offers German, English and
French through a three-position language slider; German is the initial default.
Select English to follow the interface labels used in this guide. The moon/sun
button next to the language slider switches between light and dark themes. Language
selection changes interface text, not participant contributions or topic content.

## Start the application

```{r launch, eval=FALSE}
run_app()
# Participant: http://localhost:3838
# Moderator:   http://localhost:3838/?mod=1
```

The default moderator PIN is `635`, suitable for a local demonstration. For a
classroom deployment, supply a private PIN and the externally reachable URL:

```{r classroom, eval=FALSE}
run_app(
  db_path = "/srv/brainwriting/data/classroom.sqlite",
  mod_pin = Sys.getenv("MOD_PIN"),
  base_url = "https://brainwriting.example.de",
  port = 3838,
  host = "0.0.0.0",
  poll_ms = 2500
)
```

`DB_PATH`, `MOD_PIN`, `BASE_URL` and `PORT` provide defaults when corresponding
arguments are omitted. Explicit arguments take precedence. Configuration is
resolved at launch, with no database or directory creation during package loading.
The default database is `data/brainwriting.sqlite`, relative to the working
directory. Use an absolute path for a long-running service.

The QR code of the standard session uses `base_url` exactly. A phone's `localhost`
refers to that phone; use a URL that devices in the classroom can reach. The
moderator route adds `?mod=1`; do not put this query string into the participant URL.

## Manage several sessions

One instance can hold and run several sessions at once. The bare `base_url` always
shows the standard session. Further sessions receive a short code and their own
address, for example `https://brainwriting.example.de/?s=k7m3pq`, with a matching
QR code. After signing in, **All sessions** in the moderator bar opens the overview
at `?mod=1&view=sessions`.

The overview shows each session's state, mode, participant and contribution counts,
topics and address. **New session** creates an empty session and opens its setup.
**QR code** shows the address as a large QR code that can be downloaded as PNG.
**Restart** creates a new session with the same settings and opens its setup for
review; the original results stay untouched. **Archive** moves a finished session
out of the active list: it refuses participants but remains readable, with all
exports. **Restore** returns it. **Delete** permanently removes a session and its
database file after confirmation, so export first.

The standard session keeps the bare address and is therefore never deleted or
archived in place. **Reset** clears it after confirmation. **Archive** stores its
finished results as a new archived session and then clears it for the next activity.

The PIN is entered once per browser tab, and a login lasts 12 hours, also in tabs
that stay open. The server keeps a login token in memory
and the tab keeps it in session storage, so opening other sessions does not ask
again; **Sign out** ends it and a server restart signs everyone out. Five wrong
PINs pause new logins for 15 seconds, doubling up to five minutes. Use a long,
private PIN, because it now protects deletion as well.

## Choose a device mode

| Devices | Mode | Authors | Schedule |
|:--|:--|:--|:--|
| Everyone has a phone or laptop | `individual` | People | Parallel timed rounds |
| One device per group table | `group_device` | Groups | Parallel timed rounds |
| One computer for the entire room | `hot_seat` | People | Sequential timed turns within passes |

The mode is selected during setup and cannot change during the session. All
three modes use the same rotation and entry-storage engine.

### Individual devices

The default workflow is unchanged. Prepare 2–6 topics, open the lobby and share
the QR code. People enter a name or pseudonym and join. A prepared roster adds
one-tap name buttons; claimed names are disabled, with free-text joining still
available. A blank free-text name receives an automatically generated label.
Start requires at least as many people as topics.

The app randomizes balanced groups. With 16 people and three topics, the sizes
are 6, 5 and 5. People use their own devices, read the earlier contributions on
their assigned sheet and answer both questions. Submit marks both answers as
submitted; they remain editable until the round ends. The moderator sees group
progress, can add 60 seconds and can end a round early.

### Group devices

Each device claims one named group, which becomes the author of its answers.
There is exactly one sheet per topic. Start normally requires every group to
be claimed; the moderator can explicitly start without all groups.

Reclaiming a group asks for confirmation and reuses the same persistent author
identity. Existing contributions retain their attribution. Close the old device's
tab after taking over: simultaneous edits from old and new devices use
last-write-wins. Unclaimed groups may join after the session has started.

### Hot seat

Enter at least one name in the roster, one per line. Saving starts directly at a
handover screen instead of opening a QR lobby. The next person selects the start
button to arm the timer. Finishing saves both answers and hands over to the next
person; skipping a handover creates no entries. Turn order follows roster arrival
order. A pass visits everyone once, then the topic rotation advances.

Virtual groups use the same balanced assignment as individual mode. If there are
fewer people than topics, an empty starting group receives one empty sheet so
that the shared modulo mapping remains valid. Other sheet counts remain frozen.
The estimated working duration is `passes * people * turn_secs`; it excludes
handovers and pauses. The default turn is 90 seconds.

Keep a PIN-authenticated moderator tab open on the same computer to add 30 seconds,
skip a person, end the rest of a pass, append a latecomer or finish early while
retaining results. Latecomers join the smallest virtual group and receive turns
at the end of the current and remaining passes. The kiosk derives identity from
the active turn and does not require participant identity in browser storage.

## Try the example question set

Select **Use example questions** in setup to populate three ready-to-use classroom
topics in the currently selected interface language. They follow the 6-3-5
principle: one open problem per topic, up to three new ideas in the first field,
and building on an idea from the sheet in the second. The questions remain editable.
Clearing the checkbox restores the previous topic count and questions. The option
preserves the selected device mode, roster and timers. A valid YAML upload replaces
the example selection. Switching language does not translate existing questions;
choose the interface language before enabling the examples.

## Reuse session settings

The setup form and YAML upload prepare the same fields. Upload validates and
prefills the form; it never starts a session. Review the values before saving.
Settings export is available in setup and the lobby. For hot-seat mode, export
before starting if a separate reusable settings file is needed. A finished RDS
snapshot also contains the configuration.

```yaml
format: brainwriting635-settings/1
mode: individual
rounds: 3
round_secs: 300
turn_secs: 90
topics:
  - title: "Teamwork"
    q1: "What helps a team collaborate?"
    q2: "How can we apply this next week?"
  - title: "Learning spaces"
    q1: "What makes a space useful for learning?"
    q2: "Which change could we try first?"
  - title: "Feedback"
    q1: "What makes feedback actionable?"
    q2: "How can we build it into our routine?"
groups:
  - "North"
  - "South"
  - "West"
participants:
  - "Alex"
  - "Sam"
  - "Robin"
```

The format identifier remains `brainwriting635-settings/1` for compatibility.
Topic count must be 2–6, with nonempty titles and both questions. Rounds must be
an integer from 1–12. Parallel rounds allow 30–1800 seconds; hot-seat turns allow
20–600 seconds. Group names must be unique and match the topic count. Optional
roster names must be nonempty and unique after trimming.

The upload limit is 100 KiB. YAML expressions are not evaluated. Validation
collects all errors together; unknown keys warn and are ignored. The moderator
PIN is neither loaded from nor written to the settings file.

## Understand the rotation

Group `g` in round or pass `r` receives topic
`((g - 1 + r - 1) %% K) + 1`:

```{r rotation-table}
k <- 3L
rotation <- outer(seq_len(k), seq_len(k), function(g, r) topic_for(g, r, k))
dimnames(rotation) <- list(paste("Group", 1:k), paste("Round", 1:k))
knitr::kable(rotation)
```

Every row and column is a permutation of the topics. Across K complete rounds,
every group visits every topic once. Fewer rounds cover a partial cycle; more
rounds repeat topics. A late arrival cannot recover missed topics automatically.

Each topic begins with one sheet per person in its starting group, frozen at
start, except for the single-sheet group-device mode and empty hot-seat groups.
Within-group index `i` receives sheet `((i - 1) %% n_sheets) + 1`:

```{r uneven}
sheet_for(idx = 1:6, n_sheets = 5)
sheet_for(idx = 1:5, n_sheets = 6)
```

Six people visiting five sheets share one sheet, retaining both authors' entries.
Five people visiting six sheets leave one untouched in that round. Unequal group
sizes therefore do not guarantee a contribution to every sheet every round.
Completion of either question is voluntary.

## Weigh contributions in the plenum

After the last round, the finished moderator view offers **Analyse & Plenum**.
**Start weighting** switches all participant devices to a weighting page: every
nonempty contribution of each topic, without names and in an order that differs
between people but stays stable for each person. Everyone has 100 % per topic and
distributes it with sliders in 5 % steps. A slider stops at the remaining budget,
and each change is saved immediately. Own contributions may be weighted as well.
Group devices weight as one author; on a shared hot-seat device the weighting passes
from person to person, and the moderator can skip someone.

While the weighting runs, the moderator sees per topic how many people have used
their full budget. Results remain hidden until **End weighting**, so the room is not
influenced by a live ranking. The overview then shows the highest-weighted
contributions per topic as a chart and the full ranking with the share of all points
given in the topic, the mean per person and the number of supporters. **Reopen**
continues with the saved weights. Treat the result as a snapshot of the group's
opinion to structure the discussion, not as a quality score.

Weights are stored with the voter's participant ID so that every person keeps one
budget. The app never shows who gave which weight, and the CSV, XLSX, Markdown and
PDF outputs are aggregated per contribution. Only the RDS snapshot keeps individual
weights linked to participant IDs; tick the pseudonymization option to replace them.
Each slider is named by its question and contribution for screen readers. If a person
has the page open twice, or a group device is taken over, all open pages follow the
stored weights within a few seconds.

## Read the results responsibly

The finished moderator screen separates the contribution list from descriptive
analytics. The four indicators count nonempty saved entries, average unfiltered
word count, submitted share and distinct filtered terms. Drafts are included.
The submitted share uses nonempty stored entries as its denominator, not the
number of theoretically possible answers or attendance.

Five shared figure functions provide contributions by topic and round, eight top
terms per topic, a co-occurrence network, a wordcloud and lexical continuity
between rounds. The network defaults to at least two co-occurring entries and
at most 40 frequent terms. The wordcloud uses at most 60 terms. A topic selector
filters both. Empty or insufficient text produces a placeholder.

Text processing splits at Unicode nonletters, lowercases using German locale,
retains words with at least three letters and removes German Snowball stopwords.
Topic-title and question vocabulary are also excluded by default; the app can
disable this filter. There is no stemming or semantic model. Interface language
selection does not change this German-language text-analysis method; interpret
English or French contributions accordingly.

Term counts retain repeated words, while network edges count the entries that
contain a pair once each. Lexical continuity is the Jaccard overlap between
consecutive rounds' token sets on each sheet and question, pooling authors who
share the sheet. Missing or empty rounds yield missing values rather than zero;
valid comparisons are averaged by topic. Word reuse is not evidence of idea
quality, understanding or conceptual development. Read the original answers
alongside the figures.

## Choose an export

| Format | Intended use | Contents |
|:--|:--|:--|
| CSV | Statistical or spreadsheet import | Long table, including empty drafts |
| Markdown | Readable discussion protocol | Questions and nonempty answers grouped by sheet |
| RDS | Lossless re-analysis in R | Raw tables incl. weights, settings, generation time and package version |
| XLSX | Workbook review | Contributions, authors, topics, descriptive summaries and weights |
| PDF | Shareable overview | Two A4 pages with the app's figures, plus a plenum page when weighted |
| Weighting (CSV) | Plenum ranking | Topic, rank, contribution, points, share, mean and supporters |

On macOS, the CRAN R build may require [XQuartz](https://www.xquartz.org/)
for Cairo graphics. Install it before using PDF reports; see the
[R Cairo device documentation](https://search.r-project.org/R/refmans/grDevices/html/cairo.html).

The PDF uses Cairo and patchwork directly, without a runtime dependency on
Pandoc or LaTeX. Its first page contains parameters, indicators and overview
charts. Its second contains networks, a wordcloud and per-topic summaries.
A third page shows the plenum weighting when weights exist.
Up to three topics receive separate networks; larger sessions use one combined
network. Export schemas and the PDF retain their German labels independently
of the interface language.

Optional author pseudonymization applies stable arrival-order codes to all five
downloads. It also replaces persistent author IDs and stored roster settings.
On-screen views retain names. This is not complete anonymization: personal
information typed into answers or topic prompts remains unchanged. Review such
content before sharing. CSV values should be imported as text, and Markdown
should be rendered with raw HTML disabled. For spreadsheet software, tick **Make CSV
safe for spreadsheets**: `=`, `+`, `-`, `@`, a tab or a carriage return at the start of
a cell or after an embedded comma, semicolon, tab or line break is prefixed with an
apostrophe. The default CSV stays raw.

## Persistence, recovery and deployment

SQLite stores the session, topics, authors and entries. Each operation opens and
closes its own connection, with WAL mode and a 5000 ms busy timeout. Entry
uniqueness includes author identity, so two people sharing a sheet do not replace
each other's contributions. Autosaves preserve an explicit submission flag.

The main database at `db_path` holds the standard session and the catalog of
further sessions. Each further session is a separate SQLite file with the same
schema in a `sessions/` folder next to it; deleting a session removes its file.
Existing databases migrate in place by adding `mode`, `current_turn`,
`settings_yaml`, `plenum` and `plenum_turn` and a `votes` table; `run_app()` migrates the
main database and every catalogued session file at startup, and the existing session
becomes the standard session. Older sessions retain individual mode. The settings column
preserves prepared names and turn duration across restarts; no sidecar file is
needed. Back up before upgrading and avoid operating newer-mode databases with
an older application version.

Every connected client polls the authoritative deadline. Guarded updates prevent
racing clock ticks from advancing twice. Hot-seat updates also guard the current
turn, and stale repeated controls cannot advance a later person's turn. Without
any connected clients there is no scheduler: the next connection resumes the
expired transition, without replaying every missed round.

A disconnect triggers a reload after 1.5 seconds. Individual and group devices
resend the identity stored for that session; the server resumes it or clears it
after reset. One device can therefore take part in several sessions.
Keep the same browser and origin. With browser storage disabled or cleared,
automatic identity recovery is unavailable. There is no offline editing queue.

Typing is debounced for 1200 milliseconds. Main answer fields are replaced only
on genuine session, round or assignment transitions, not every poll or language
change. A delayed save retains its original editing context. Keystrokes in the
last 1–2 seconds of a round or turn can still lose the race with its boundary.
Submit before expiry. The timer turns red below one minute and can briefly show
zero until the next poll.

Use exactly one R process per data folder, a writable local disk and a persistent
database directory. Do not use replicas, ShinyProxy or horizontal scaling.
Parallel courses can share one instance as separate sessions.
The supplied Docker image runs as UID 10001. Compose expects an existing Traefik
installation, external `proxy` network, TLS resolver and reachable hostname.

```{bash deploy, eval=FALSE}
mkdir -p docker/data
sudo chown 10001:10001 docker/data
export BW_HOST=brainwriting.example.de
export TRAEFIK_CERTRESOLVER=letsencrypt
export MOD_PIN='replace-with-a-private-pin'
docker compose -f docker/docker-compose.yml up -d --build
```

Mount the whole database directory, including SQLite WAL files and the
`sessions/` folder. For backups,
stop the app before copying it or use SQLite's online backup API. A live database
file copied without its WAL can omit committed changes. Restarting with the same
volume resumes the session; rebuilding an image does not reset it.

Export before resetting a finished session. Confirmed reset removes logical
session data and clears participant identity on reconnect. It does not remove
old downloads, backups, snapshots or residual SQLite pages. Apply a retention
policy to those copies as well. Self-hosting and pseudonyms alone do not establish
legal compliance.

## Rehearse and troubleshoot

| Symptom | Check |
|:--|:--|
| Phone cannot open the QR link | Public URL, DNS, Wi-Fi reachability and TLS |
| Database cannot be created | Mounted directory ownership for UID 10001 |
| Repeated reconnects | Proxy WebSocket support and service availability |
| Duplicate participant | Cleared storage or a separate browser profile |
| Timer waited while everyone was absent | Expected: no clients means no polling scheduler |
| A sheet lacks a round | Unequal groups, late arrivals or unanswered questions |
| Old work returns after restart | Expected persistence; archive it or start a new session |
| A session link shows “Session nicht gefunden” | The code is mistyped, or the session was deleted |
| Logins pause after wrong PINs | Expected throttling; wait for the shown time |

Tests cover rotation, persistence, configuration, mode transitions, reconnection,
exports, report page counts, session management and browser workflows. A passing automated suite does
not replace a rehearsal on the institution's actual devices and network.
